A mechanization of semi-simplicial and semi-cubical sets in the calculus of inductive constructions

### 2022-11-09 paper/paper.tex
paper: minor polish

To assign $X_0, X_1, X_2, \ldots$ in the above indexed representation types, we need to exhibit a recursive definition over our building blocks, which we will call $\framep$, $\layer$, and $\filler$. A $\framep$ is a boundary of a standard form (simplex, cube, etc.), a $\filler$ is a filled $\framep$.

As a first step, we assign every $X_n$ the type $\fullframe[n] \rightarrow \U$ uniformly, applying the isomorphism between $A \rightarrow B \rightarrow C$ and $A \times B \rightarrow C$, or between $\Pi a : A, (B a \rightarrow C)$ and $(\Sigma a : A, B a) \rightarrow C$. Let us illustrate how we will build $\fullframe[n]$ by refining types of $X_0, X_1, \ldots, X_{n - 1}$ so that they have the form $\fullframe[n] \rightarrow \U$. To begin, let us set $\fullframe[0] = \unittype$, so that the type $\U$ of $X_0$ can be equivalently formulated as $\unittype \rightarrow \U$. The rest of the construction follows.

\begin{equation*}
  \begin{array}{llll}
    X_0                                :                & \underbrace{\unittype}_{\framep[0][0]}                                                   & \rightarrow \U \\
    X_1                                :                & \underbrace{\Sigma \unitpoint: \unittype. \underbrace{\left(
    \begin{array}{c}
        \underbrace{X_0(\unitpoint)}_{\filler[0][0]}
        \\ \times \\
        \underbrace{X_0(\unitpoint)}_{\filler[0][0]}
      \end{array}\right)}_{\layer[1][0]}}_{\framep[1][1]} & \rightarrow \U                                                                                            \\
    X_2                                :                & \underbrace{\Sigma a: \underbrace{\left(\Sigma \unitpoint: \unittype. \underbrace{\left(
        \begin{array}{c}
          \underbrace{\Sigma b: \left(
          \begin{array}{c}
              X_0(\unitpoint)
              \\ \times \\
              X_0(\unitpoint)
            \end{array}
          \right).\; \underbrace{X_1 \underbrace{(\unitpoint, b)}_{\restrf[2][0][L]}}_{\filler[1][1]}}_{\filler[1][0]}
          \\ \times \\
          \underbrace{\Sigma b: \left(
          \begin{array}{c}
              X_0(\unitpoint)
              \\ \times \\
              X_0(\unitpoint)
            \end{array}
          \right).\; \underbrace{X_1 \underbrace{(\unitpoint, b)}_{\restrf[2][0][R]}}_{\filler[1][1]}}_{\filler[1][0]}
        \end{array}
        \right)}_{\layer[2][0]}\right)}_{\framep[2][1]}.\underbrace{\left(
      \begin{array}{c}
        \underbrace{X_1 \underbrace{\left(a.\hd, \left(
          \begin{array}{c}
            a.\tl.L.\hd.L, \\
            a.\tl.R.\hd.L
          \end{array}
          \right)\right)}_{\restrf[2][1][L]}}_{\filler[1][1]}
        \\ \times \\
        \underbrace{X_1 \underbrace{\left(a.\hd, \left(
          \begin{array}{c}
            a.\tl.L.\hd.R, \\
            a.\tl.R.\hd.R
          \end{array}
          \right)\right)}_{\restrf[2][1][R]}}_{\filler[1][1]}
      \end{array}
      \right)}_{\layer[2][1]}}_{\framep[2][2]}
                                                        & \rightarrow \U                                                                                            \\
    \ldots
  \end{array}
\end{equation*}

$\fullframe[n]$ is a nesting of $n$ $\Sigma$-types, of which each component is called a $\layer[n]$.

The question is then to find the recursive recipe defining $\fullframe[n]$. First, $\fullframe[n]$ depends on all $X_i$ upto $n - 1$. So, we need to package up $X_i$, for $i < n$, into a type, which we call $\X[][<n][]$. This allows us to give the type $\X[][<n][] \rightarrow \U$ to $\fullframe[n]$. Then, for $D: \X[][<n][]$, representing an initial prefix of $X_0, X_1, \ldots X_{n - 1}$, the indexed set $X_n$ has type $\fullframe[n](D) \rightarrow \U$. We shall give the full definition of $\fullframe[n]$ shortly, but first, assuming $\fullframe[n]$ is defined, we are able to take the coinductive limit of the $\X[][<n][]$, as shown in \ref{tab:coind}, therefore fully characterizing the indexed definition of $\nu$-sets.

>>>>

To assign $X_0, X_1, X_2, \ldots$ in the above indexed representation types, we need to exhibit a recursive definition over our building blocks, which we will call $\framep$, $\layer$, and $\filler$. A $\framep$ is a boundary of a standard form (simplex, cube, etc.), which we decompose into $\layer$, and a $\filler$ is a filled $\framep$.

We assign every $X_n$ the type $\fullframe[n] \rightarrow \U$ uniformly, applying the isomorphism between $A \rightarrow B \rightarrow C$ and $A \times B \rightarrow C$, or between $\Pi a : A, (B a \rightarrow C)$ and $(\Sigma a : A, B a) \rightarrow C$. Let us illustrate how we build $\fullframe[n]$ by refining types of $X_0, X_1, \ldots, X_{n - 1}$ so that they have the form $\fullframe[n] \rightarrow \U$. To begin, let us set $\fullframe[0] = \unittype$, so that the type $\U$ of $X_0$ can be equivalently formulated as $\unittype \rightarrow \U$. Each $\fullframe[n]$ is made of $n$-layers, written $\layer[n][p]$ with $p < n$, initalized with the $\unittype$ type. Then, $X_1$ is made of one layer, so that it can be written as a $\Sigma$-type of an inhabitant of the $\unittype$ and $\layer[1][0]$. The rest of the construction follows.

\begin{equation*}
  \begin{array}{ll}
    X_0                                : \underbrace{\unittype}_{\framep[0][0]}                                                    \rightarrow \U \\
    X_1                                : \underbrace{\Sigma \unitpoint: \unittype. \underbrace{\left(
    \begin{array}{c}
        \underbrace{X_0(\unitpoint)}_{\filler[0][0]}
        \\ \times \\
        \underbrace{X_0(\unitpoint)}_{\filler[0][0]}
      \end{array}\right)}_{\layer[1][0]}}_{\framep[1][1]} \rightarrow \U                                                                            \\
    X_2                                : \underbrace{\Sigma a: \underbrace{\left(\Sigma \unitpoint: \unittype. \underbrace{\left(
        \begin{array}{c}
          \underbrace{\Sigma b: \left(
          \hspace{-0.4em}\begin{array}{c}
                           X_0(\unitpoint)
                           \\ \times \\
                           X_0(\unitpoint)
                         \end{array}\hspace{-0.4em}
          \right).\; \underbrace{X_1 \underbrace{(\unitpoint, b)}_{\restrf[2][0][L]}}_{\filler[1][1]}}_{\filler[1][0]}
          \\ \times \\
          \underbrace{\Sigma b: \left(
          \hspace{-0.4em}\begin{array}{c}
                           X_0(\unitpoint)
                           \\ \times \\
                           X_0(\unitpoint)
                         \end{array}\hspace{-0.4em}
          \right).\; \underbrace{X_1 \underbrace{(\unitpoint, b)}_{\restrf[2][0][R]}}_{\filler[1][1]}}_{\filler[1][0]}
        \end{array}
        \right)}_{\layer[2][0]}\right)}_{\framep[2][1]}.\underbrace{\left(
      \begin{array}{c}
        \underbrace{X_1 \underbrace{\left(a.\hd, \left(
            \hspace{-0.4em}\begin{array}{c}
                             a.\tl.L.\hd.L, \\
                             a.\tl.R.\hd.L
                           \end{array}\hspace{-0.4em}
            \right)\right)}_{\restrf[2][1][L]}}_{\filler[1][1]}
        \\ \times \\
        \underbrace{X_1 \underbrace{\left(a.\hd, \left(
            \hspace{-0.4em}\begin{array}{c}
                             a.\tl.L.\hd.R, \\
                             a.\tl.R.\hd.R
                           \end{array}\hspace{-0.4em}
            \right)\right)}_{\restrf[2][1][R]}}_{\filler[1][1]}
      \end{array}
      \right)}_{\layer[2][1]}}_{\framep[2][2]}
    \rightarrow \U                                                                                                                                \\
    \ldots
  \end{array}
\end{equation*}

Let us then make explicit the recursive recipe defining $\fullframe[n]$. First, $\fullframe[n]$ depends on all $X_i$ upto $n - 1$. So, we need to package up $X_i$, for $i < n$, into a type, which we call $\X[][<n][]$. This allows us to give the type $\X[][<n][] \rightarrow \U$ to $\fullframe[n]$. Then, for $D: \X[][<n][]$, representing an initial prefix of $X_0, X_1, \ldots X_{n - 1}$, the indexed set $X_n$ has type $\fullframe[n](D) \rightarrow \U$. We shall give the full definition of $\fullframe[n]$ shortly, but first, assuming $\fullframe[n]$ is defined, we are able to take the coinductive limit of the $\X[][<n][]$, as shown in \ref{tab:coind}, therefore fully characterizing the indexed definition of $\nu$-sets.
